Manuscript: Detecting differences in Size Spectra

### 2022-09-24 Manuscript.tex
new ms knit. Added pub estimates to SI

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Manuscript},
  pdfauthor={Justin Pomeranz1,; James R. Junker2,3; Vojsava Gjoni4; Jeff S. Wesner4},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{indentfirst}
\linenumbers
\newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Manuscript}
\author{Justin Pomeranz\textsuperscript{1,*} \and James R.
Junker\textsuperscript{2,3} \and Vojsava
Gjoni\textsuperscript{4} \and Jeff S. Wesner\textsuperscript{4}}
\date{05 September, 2022}

@@

@@

Alternate plot version

\begin{figure}
\centering
\includegraphics{figures/PLB_main_v2.png}
\caption{alternate plot version, thoughts on which you prefer??}
\end{figure}

All methods performed reasonably well in detecting the known
relationship across the hypothetical environmental gradient (mean
\(\beta_1\) estimates ranged from -0.419 to -0.498). However, the SD for
the NAS and ELBn methods were 2-3x larger that that of the MLE method
(Figure \textsubscript{XX}). This result was consistent across
simulation parameters, including sample size, number of sites, and
\(\lambda\) values.

@@

\begin{verbatim}
##   Model   p50   p25  p975
## 1   MLE -0.49 -0.50 -0.48
## 2  ELBn -0.49 -0.52 -0.29
## 3   NAS -0.44 -0.46 -0.30
\end{verbatim}

\begin{figure}
\centering
\includegraphics{figures/n_vary_main.png}
\caption{Individual regression estimates across the hypothetical
gradient based on sample size (rows) and methodology used (columns).}
\end{figure}

@@

All of the methods performed similarly when there was no relationship
across the hypothetical gradient. The mean \(\beta_1\) estimates for all
three methods ranged from -0.0005 to 0.0002. However, the distribution
of coefficient estimates was much wider for the two binning methods; the
95th percentile of the MLE distributions is narrower thatn the 66th
percentiles for the NAS and ELBn methods.

@@

\begin{verbatim}
##   Model p50   p25 p975
## 1   MLE   0 -0.01 0.04
## 2   NAS   0 -0.04 0.12
## 3  ELBn   0 -0.05 0.13
\end{verbatim}

@@

The performance of all methods to detect small variation in the
\(\lambda\) parameter (between -1.9 to -2.1) across the environmental
gradient was also investigated. The median \(\beta_1\) coefficient
estimates for the MLE method was -0.1, which exactly matched the value
for the hypothetical relationship, while the other two methods both had
median estimates of 0.9. The upper 97.5 percentile of the distribution
was below 0 for the MLE method, meaning that most of the estimates had
the correct sign. In contrast, the upper 97.5 percentile for both the
NAS and ELBn methods were greater than 0 at 0.027 and 0.052,
respectively. (Figure 4).

@@

\begin{verbatim}
##   Model   p50   p25  p975
## 1   MLE -0.10 -0.11 -0.06
## 2   NAS -0.09 -0.13  0.03
## 3  ELBn -0.09 -0.13  0.05
\end{verbatim}

@@

\textbf{NOTE} I need to re-plot these in ggplot and and stitich
together. Currently stitching pre-made png files, I think that's why it
looks so funky.

@@

\begin{figure}
\centering
\includegraphics{figures/empirical_combined.png}
\caption{Figure of empirical data estimates. All of the methods estimate
the same sign of the relationship, but the estimates from the binning
methods are generally greater than the MLE estimates.}
\end{figure}

@@

The \(\beta_1\) coefficient estimates for the AMD data are all positive
and of relatively similar magnitudes across methods. The MLE estimate
has the smallest magnitude and SD (0.063 \(\pm\) 0.011 SD), The ELBn
method is slightly larger and has a wider SD : (0.072 \(\pm\) 0.018).
Finally, the NAS method has the largest estimate and an intermediate
standard deviation (0.076 \(\pm\) 0.013). The range of the gradient in
the AMD data is relatively small (9.5), meaning that the absolute change
in the size spectra parameter (\(\lambda\)) across the AMD gradient are
similar regardless of method used (range of absolute: 0.59 to 0.72).

Similar to the AMD results, the \(\beta_1\) coefficient estimates for
the NEON data are all negative and have similar magnitudes across
methods. The MLE estimate is once again the smallest (-0.006 \(\pm\)
0.001 SD). The NAS and ELBn methods are larger at -0.004 \(\pm\) 0.002,
and -0.002 \(\pm\) 0.002, respectively. Although the estimates are
similar and have a small magnitude, the range of the gradient in this
data is larger (\textasciitilde30\(\circ\)C), meaning that the absolute
change in \(\lambda\) estimates ranges from 0.06 to 0.17, depending on
the method used.

@@

Estimates of the relationship between lambda and a hypothetical gradient
varied depending on the method used. However, interpretations of
empirical data were broadly consistent across methodologies. Upon closer
inspection, the values of size spectra parameters in the empirical data
were considerably shallower than -2. Given that we found the performance
of all methods increased with shallower \(\lambda\)'s we wanted to
investigate how the methods performed with simulated values closer to
the empirical estimates of parameters describing size spectra
relationships. Therefore, we repeated the simulation process as in the
main analysis, but used \(\lambda\) values ranging from -1.1 to -1.5,
with a known relationship of \(\beta_1 = -0.2\)

@@

\begin{figure}
\centering
\includegraphics{figures/shallow_lambda_main.png}
\caption{Relationship estimates across the hypothetical gradient for
each replicate. Each panel is a different method for estimating the size
spectra parameter.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{figures/PLB_shallow_lambda_relationship_density.png}
\caption{Distribution of relationship coefficient estimates. Vertical
line is the known relationship. All methods under estimate the value,
but the mean magnitude and distribution of values is greater for the
ELBn and NAS methods.}
\end{figure}

\begin{verbatim}
##   Model   p50   p25  p975
## 1   MLE -0.20 -0.21 -0.18
## 2  ELBn -0.19 -0.21 -0.16
## 3   NAS -0.17 -0.18 -0.15
\end{verbatim}

@@

Generally, the estimate from the MLE method was always closer to the
known relationship in simulated data, and always had the smallest
variation around the estimate.

Binning methods are easy to use and interpret, which most likely
accounts for their wide use in ecological studies. However, aggregating
individuals into logarithmic bins removes a large source of the
variation within the data by collapsing body size variation into a
single values within bins. For example, all individuals placed into a
bin that ranges from 2-4 grams of mass are all treated as having a mass
of 3 grams, the midpoint of that bin. Likewise, a single abundance value
is taken for this bin, despite that fact that there is almost certainly
variation in the abundance of individuals that weigh \textasciitilde2
grams or \textasciitilde4 grams. By homogenizing the data in this way,
it seems that binning methods will always produce noisier results than
MLE from the binning process alone, since binning is akin to deleting
information that the model could otherwise use.

One of the benefits of using MLE is that all data points are retained
for use within the model.

@@

Despite the drop in performance with reduced variation in \(\lambda\)
values, when a significant relationship was observed, it was generally
in the correct direction and of a similar magnitude. This suggests that
previously reported significant changes in size spectra parameters
across environmental gradients and in experimental manipulations are
plausible, and the magnitude of the relative change is a reasonable
estimate. Given that all of the data within a study is treated
identically, the the over all change in size spectra parameters is
likely reasonable. However, the biases and inconsistencies in
relationship estimates presented here suggest that it would be difficult
if not impossible to compare the relative changes across different
published studies which use different methods. Publication of individual
body size data with future studies of size spectra relationships would
greatly aid in our ability to generalize changes to this fundamental
aspect of community organization across spatiotemporal scales and in
response to environmental stressors and perturbations.

@@

Both of the binning methods can perform as well as MLE under certain
circumstances (large sample size, shallow \(\lambda\)). However, it
would be extremely difficult to know if this was the case beforehand.
Furthermore, in order to confirm this result, MLE methods would likely
need to be employed. With the publication of the sizespectra package
(2017), producing MLE estimates of size spectra parameters is a
relatively easy task. Therefore, we recommend using this in all future
studies of size spectra relationships.

@@

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Manuscript: Detecting differences in Size Spectra},
  pdfauthor={Justin Pomeranz1,; James R. Junker2,3; Vojsava Gjoni4; Jeff S. Wesner4},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{indentfirst}
\linenumbers
\newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Manuscript: Detecting differences in Size Spectra}
\author{Justin Pomeranz\textsuperscript{1,*} \and James R.
Junker\textsuperscript{2,3} \and Vojsava
Gjoni\textsuperscript{4} \and Jeff S. Wesner\textsuperscript{4}}
\date{24 September, 2022}

@@

\hypertarget{data-availabity}{%
\subsection{Data Availabity}\label{data-availabity}}

The data used in this manuscript is already publicly available (Pomeranz
et al.~2019 data dryad DOI \textbf{here}, NEON data citation
\textbf{here}. R Scripts to reproduce the full simulation and analysis
are available at:
({[}\url{https://github.com/Jpomz/detecting-spectra-differences}{]}).

@@

\begin{figure}
\centering
\includegraphics{figures/PLB_main_v2.png}
\caption{alternate plot version; I think I like this one better. What
are your thoughts on which you prefer??}
\end{figure}

The distribution of \(\beta_1\) estimates from MLE method were the most
accurate and precise, and the NAS method was the least accurate and
precise (Table \emph{xx}). The median estimates for all methods were
closer to the known relationship. All of the methods produced estimate
distributions with long right tails, but this pattern was greater in the
ELBn and NAS methods (e.g., SD for ELBn and NAS methods were 2-3x
larger, Figure \emph{xx} Halfeye plots).This result was consistent
across simulation parameters, including sample size, number of sites,
and \(\lambda\) values.

@@

\begin{verbatim}
##   Model   mean    p50    p25   p975
## 1   MLE -0.495 -0.494 -0.499 -0.483
## 2  ELBn -0.460 -0.489 -0.524 -0.291
## 3   NAS -0.419 -0.442 -0.458 -0.302
\end{verbatim}

\begin{figure}
\centering
\includegraphics{figures/n_vary_main.png}
\caption{Individual regression estimates across the hypothetical
gradient based on sample size (rows) and methodology used (columns).
(match this figure to ``new'' style if we like that better)}
\end{figure}

@@

All of the methods performed similarly when there was no relationship
across the hypothetical gradient. The mean \(\beta_1\) estimates for all
three methods ranged from -0.0005 to 0.0002. Although the distributions
were closely centered on 0, the distribution of coefficient estimates
was much wider for the two binning methods; the 95th percentile of the
MLE distributions is narrower than the 66th percentiles for the NAS and
ELBn methods.

@@

\begin{verbatim}
##   Model   mean    p50    p25  p975
## 1   MLE -0.001 -0.001 -0.012 0.038
## 2   NAS  0.000 -0.001 -0.039 0.119
## 3  ELBn  0.000 -0.001 -0.046 0.134
\end{verbatim}

@@

The performance of all methods to detect small variation in the
\(\lambda\) parameter (between -1.9 to -2.1) across the environmental
gradient was also investigated. The median \(\beta_1\) coefficient
estimates for the MLE method was -0.1, which exactly matched the value
for the hypothetical relationship, while the other two methods both had
median estimates of -0.09. The upper 97.5 percentile of the distribution
was below 0 for the MLE method, meaning that most of the estimates had
the correct sign. In contrast, the upper 97.5 percentile for both the
NAS and ELBn methods were greater than 0 at 0.027 and 0.052,
respectively. (Figure 4).

@@

\begin{verbatim}
##   Model   mean    p50    p25   p975
## 1   MLE -0.100 -0.101 -0.112 -0.062
## 2  ELBn -0.091 -0.095 -0.134  0.052
## 3   NAS -0.088 -0.091 -0.126  0.027
\end{verbatim}

@@

@@

\begin{figure}
\centering
\includegraphics{figures/empirical_combined.png}
\caption{Figure of empirical data estimates. All of the methods estimate
the same sign of the relationship, but the estimates from the binning
methods are generally greater than the MLE estimates. \textbf{NOTE} I
need to re-plot these in ggplot and and stitch together. Currently
stitching pre-made png files, I think that's why it looks so funky.}
\end{figure}

@@

The \(\beta_1\) coefficient estimates for the AMD data are all positive
and of relatively similar magnitudes across methods. The MLE estimate
has the smallest magnitude and SD (0.063 \(\pm\) 0.011 SD), The ELBn
method is slightly larger and has a wider SD : (0.072 \(\pm\) 0.018).
Finally, the NAS method has the largest estimate and an intermediate
standard deviation (0.076 \(\pm\) 0.013). The range of the gradient in
the AMD data is relatively small (9.5), but the absolute change in the
size spectra parameter (\(\lambda\)) across the AMD gradient vary by
\textasciitilde0.13 units depending on the method used (range of
absolute change: 0.59 to 0.72).

Similar to the AMD results, the \(\beta_1\) coefficient estimates for
the NEON data are all negative and have similar magnitudes across
methods. The MLE estimate is once again the smallest (-0.006 \(\pm\)
0.001 SD). The NAS and ELBn methods are larger at -0.004 \(\pm\) 0.002,
and -0.002 \(\pm\) 0.002, respectively. Although the estimates are
similar and have a small magnitude, the range of the gradient in this
data is larger (\textasciitilde30\(\circ\)C), meaning that the absolute
change in \(\lambda\) estimates varies by 0.11 units depending on the
methods used (range of absolute change across gradient: 0.06 to 0.17).

@@

Estimates of the relationship between lambda and a hypothetical gradient
varied depending on the method used. However, interpretations of
empirical data were broadly consistent across methodologies. Upon closer
inspection, the values of size spectra parameters in the empirical data
were considerably shallower than -2. Given that we found the performance
of all methods increased with shallower \(\lambda\)'s we wanted to
investigate how the methods performed with simulated values closer to
the empirical estimates of parameters describing size spectra
relationships in streams. Therefore, we repeated the simulation process
as in the main analysis, but used \(\lambda\) values ranging from -1.1
to -1.5, with a known relationship of \(\beta_1 = -0.2\)

@@

\begin{figure}
\centering
\includegraphics{figures/shallow_lambda_main.png}
\caption{Relationship estimates across the hypothetical gradient for
each replicate. Each panel is a different method for estimating the size
spectra parameter. (match to ``new'' style if needed)}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{figures/PLB_shallow_lambda_relationship_density.png}
\caption{Distribution of relationship coefficient estimates. Vertical
line is the known relationship. All methods under estimate the value,
but the mean magnitude and distribution of values is greater for the
ELBn and NAS methods. (not sure why this is such different proportions,
need to re plot with right dimensions)}
\end{figure}

\begin{verbatim}
##   Model   mean    p50    p25   p975
## 1   MLE -0.199 -0.199 -0.205 -0.183
## 2  ELBn -0.195 -0.195 -0.207 -0.161
## 3   NAS -0.174 -0.174 -0.183 -0.151
\end{verbatim}

\textbf{NOTE} Not sure if these tables are usful or not. Shows same
information as figures, but might be easier to see the quantitative
values instead of estimating from plot

@@

Generally, the estimate from the MLE method was always closer to the
known \(\lambda\) value, and resulted in the most accurate \(\beta_1\)
coefficient estimate for the hypothetical relationship in simulated
data, and always had the smallest variation around the estimate.

Binning methods are easy to use and interpret, which most likely
accounts for their wide use in ecological studies. However, aggregating
individuals into logarithmic bins removes a large source of the
variation within the data by collapsing body size variation into a
single values within bins. For example, all individuals placed into a
bin that ranges from 2-4 grams of mass are all treated as having a mass
of 3 grams, the midpoint of that bin. Likewise, a single abundance value
is taken for this bin, despite that fact that there is almost certainly
variation in the abundance of individuals that weigh \textasciitilde2
grams or \textasciitilde4 grams. By homogenizing the data in this way,
it seems that binning methods will always produce noisier results than
MLE from the binning process alone, since binning is akin to deleting
information that the model could otherwise use. One of the benefits of
using MLE is that all data points are retained for use within the model.

The variation in estimated relationships across the empirical datasets
varied by \textasciitilde0.1 units depending on the method used. At
first glance this may not appear to be a concerning magnitude of
variation. However, when one considers the range of variation reported
in many seminal works on size spectra relationships in streams is
\textasciitilde0.1-0.2 (SI Table \emph{xx}), it is apparent that many of
these results may be dependent on the method used.

@@

Despite the drop in performance with reduced variation in \(\lambda\)
values, the estimated relationship coefficient was generally in the
correct direction and of a similar magnitude. This suggests that
previously reported significant changes in size spectra parameters
across environmental gradients and in experimental manipulations are
plausible. Given that all of the data within a study is treated
identically, the the over all change in size spectra parameters is
likely reasonable. However, the biases and inconsistencies in
relationship estimates presented here suggest that it would be difficult
if not impossible to compare the relative changes across different
published studies which use different methods. Publication of individual
body size data with future studies of size spectra relationships would
greatly aid in our ability to generalize changes to this fundamental
aspect of community organization across spatiotemporal scales and in
response to environmental conditions.

\textbf{NOTE} not sure if the idea below belongs in this paper or not,
but it's something I've been thinking about for a while. I'm using this
as an excuse to write out my thoughts and to see what y'all think about
it. Maybe it's another paper to write in the future?

An interesting finding is that the empirical data sets used here were
quite a bit shallower than is expected based on theory, and that what
has been observed empirically in marine systems. There are two,
non-mutually exclusive, possible explanations for this: 1) freshwater
streams are more dependent on detrital resources, 2) streams are
effectively 2D systems, whereas marine systems are 3D. Blanchard et
al.~(2009) observed that benthic marine communities had shallower slopes
than pelagic communities, and that this was likely due to a stronger
reliance on a shared detrital basal resource in the benthic community,
whereas the pelagic community was largely driven by energy transfer
through predator-prey interactions. The importance of allochthonous
resources in stream ecosystems has long been recognized (Vannote et
al.~1980). The stream data here largely comes from low order, headwater
streams which may be reasonably considered to be driven by allochthonous
resources. It is possible that the communities sampled here are
structured based on a shared basal resource, and that the importance of
predator-prey interactions is reduced. An alternate hypothesis is the
relative dimensionality of search space in benthic and pelagic systems.
Consumption rates and interaction strenghts have been shown to vary
markedly based on search dimensionality. Consumers generally encounter
resources more frequently in 3D environments (i.e., aerial, pelagic)
than in 2D environments (i.e., benthic, terrestrial), and lead to
stronger consumption rates and interaction strengths. Likewise, it has
been hypothesized that the scaling exponent for mass-specific search
rates may vary based on search space dimensionality instead of being set
at approximately 0.75 (Pawar, Dell, Savage, 2012).

@@

Both of the binning methods can perform as well as MLE under certain
circumstances (large sample size, shallow \(\lambda\)). However, it
would be unlikely for a researcher to know if this was the case
beforehand. Furthermore, in order to confirm this result, MLE methods
would likely need to be employed. With the publication of the
\emph{sizespectra} package (2017), producing MLE estimates of size
spectra parameters is a relatively easy task. Therefore, we recommend
using this in all future studies of size spectra relationships.

@@

\textbf{NOTE} udate all plots to be consistent with main text

@@

\hypertarget{comparison-with-other-published-estimates}{%
\section{Comparison with other published
estimates}\label{comparison-with-other-published-estimates}}

SI Table. This table shows published estimates of the variation in size
spectra slopes (or exponents) in empirical studies. It is unclear how to
directly compare estimates of the slope with different methods.However,
the published estimates here range from \textasciitilde0.1 to 0.2 across
the gradients studied. For comparison, the 2.5-95\% quantiles around the
relationship estimate for the MLE method were \textasciitilde0.1,
whereas for the ELBn and NAS method they were \textasciitilde0.25 and
\textasciitilde0.2, respectively. b\_diff is the change in estimate
(b-low - b-min). System refers to stream communities or mesocosm
experiments. Method: MLE = maximum likelihod estimate, ELB = equal
logarthmic binning, the number before indicates the number of bins used.
The normalization process shifts the estimates by an absolute value of
1.0. Hence, direct comparison the relative change in normalized and
non-normalized studies should not introduce any bias. The O'Gorman et
al.~2017 study used average species size and abundance (Local Size
Density Relationship, \emph{sensu} White et al.~2007) as opposed to
individual size distribution. These methods are related, but it is
unclear how to directly compare estimates from each method.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{read.csv}\NormalTok{(}\StringTok{"lit comparison.csv"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                                      Author b_diff error error_type b_low
## 1                      Pomeranz et al. 2021   0.12    NA            -1.37
## 2     Yvon Durocher et al. 2011 (Community)   0.13    NA            -0.79
## 3             Dossena et al. 2012 (October)   0.14 0.070         se    NA
## 4                      O'Gorman et al. 2017   0.15    NA            -0.85
## 5                      Martinez et al. 2016   0.15    NA            -1.11
## 6 Yvon Durocher et al. 2011 (Phytoplankton)   0.19    NA            -0.46
## 7                    McGarvey and Kirk 2018   0.19    NA            -1.81
## 8               Dossena et al. 2012 (April)   0.21 0.064         se    NA
##   b_high       System      Driver
## 1  -1.25      Streams Temperature
## 2  -0.92 FW mesocosms Temperature
## 3     NA FW mesocosms Temperature
## 4  -0.70      Streams Temperature
## 5  -0.96      Streams    Land Use
## 6  -0.65 FW mesocosms Temperature
## 7  -1.62      Streams Seasonality
## 8     NA FW mesocosms Temperature
##                                                Method
## 1                                                 MLE
## 2                              10 ELB, not normalized
## 3                               6 ELB, not normalized
## 4 ln (mean species abundance) ~ ln(mean species mass)
## 5                               6 ELB, not normalized
## 6                              10 ELB, not normalized
## 7                                      NAS, Log2 bins
## 8                               6 ELB, not normalized
\end{verbatim}
